Multilevel computation of glacier geometry \\ from Stokes dynamics

### 2021-06-06 paper/msg.tex
Another Jouvet paper to cite (follows Pralong&Funk).

However, as a consequence of the no-overhangs assumption, the SIGP as described here is likely not to be well-posed because of an issue at the ice margin.  There may be no steady state because the fluid in the vicinity of a steep ice margin, especially on a steep bed feature, ``wants'' to generate an overhang, violating the assumption that $s$ is well-defined.  Furthermore the same concern applies to each time step of an evolving model; actual glaciers can develop overhangs at the margin.  However, because overhangs are small features in large glaciers and ice sheets, essentially all modeling literature ignores this possibility and assumes well-defined surface elevation and thickness \cite{Jouvetetal2008,Lengetal2012,WirbelJarosch2020}.  An exception is \cite{PralongFunk2005}, who suggest a potentially well-posed continuum model, a nontrivial extension of the one here, in which serac and ice-cliff calving occurs via a damage variable and a stress-fracture failure criterion.  Though marginal cliffs may exist in our model (Section \ref{sec:fe}), overhangs are never allowed.

@@

To build a better smoother we return to the computation of the Jacobian diagonal entry $\rho_i'(s^h; 0)$.  Suppose we partition the support of $\psi_i$ into where there is ice and not,
\begin{equation}
\theta_i = \{\psi_i > 0\} \cap \{s^h > b^h\}, \qquad {\hat\theta}_i = \{\psi_i > 0\} \setminus \theta_i.  \label{eq:thetasupport}
\end{equation}
Recall that $\bu|_{z}$ denotes the surface velocity of the solution to the Glen-Stokes problem \eqref{eq:glenstokesweak} on the domain $\Lambda_{z}$, using an extruded mesh (Section \ref{sec:fe}).  By \eqref{eq:sigpfunctionalderiv}, \eqref{eq:idoderiv}, \eqref{eq:differencecases}, and \eqref{eq:ferhoi} we may compute
\begin{equation}
\rho_i(s^h; 0) = F^h(s^h)[\psi_i] - \ip{a}{\psi_i} = - \int_{\theta_i} (\bu|_{s^h} \cdot \bn_{s^h}- a)\, \psi_i  \label{eq:rhozero}
\end{equation}
and
\begin{align}
\rho_i'(s^h; 0) &= (F^h)'(s^h)[\psi_i,\psi_i]  \label{eq:rholocalderiv} \\
  &= - \int_{\theta_i} \lim_{\eps\to 0^+} \frac{(\bu|_{s^h+\eps\psi_i} - \bu|_{s^h}) \cdot \bn_{s^h+\eps\psi_i}}{\eps} \psi_i - \int_{{\hat\theta}_i} \lim_{\eps\to 0^+} \frac{\bu|_{s^h+\eps\psi_i} \cdot \bn_{s^h+\eps\psi_i}}{\eps} \psi_i.  \notag
\end{align}

>>>>

However, as a consequence of the no-overhangs assumption, the SIGP as described here is likely not to be well-posed because of an issue at the ice margin.  There may be no steady state because the fluid in the vicinity of a steep ice margin, especially on a steep bed feature, ``wants'' to generate an overhang, violating the assumption that the function $s$ is well-defined.  Furthermore the same concern applies to each time step of an evolving model; overhangs can develop at the margin.  Because overhangs are small features in large glaciers and ice sheets, most modeling literature ignores this possibility and assumes well-defined surface elevation and thickness \cite{Jouvetetal2008,Lengetal2012,WirbelJarosch2020}.  Exceptions \cite{PralongFunk2005} and \cite{Jouvetetal2011} provide a continuum model in which ice-cliff calving occurs via a damage variable and a stress-fracture failure criterion, and such a model might explain how (essentially) well-defined surface elevation and thickness could arise in a well-posed model.  However, such a model is a nontrivial extension of the above equations, and though marginal cliffs may exist in our model (Section \ref{sec:fe}), overhangs are never allowed herein.

@@

To build a better smoother we return to the computation of the Jacobian diagonal entry $\rho_i'(s^h; 0)$.  Suppose we partition the support of $\psi_i$ into where there is ice and not,
\begin{equation}
\theta_i = \{\psi_i > 0\} \cap \{s^h > b^h\}, \qquad {\hat\theta}_i = \{\psi_i > 0\} \setminus \theta_i.  \label{eq:thetasupport}
\end{equation}
Recall that $\bu|_{z}$ denotes the surface velocity of the solution to the Glen-Stokes problem \eqref{eq:glenstokesweak} on the domain $\Lambda_{z}$, using an extruded mesh (Section \ref{sec:fe}).  By \eqref{eq:idoderiv}, \eqref{eq:differencecases}, and \eqref{eq:ferhoi} we may compute
\begin{equation}
\rho_i(s^h; 0) = F^h(s^h)[\psi_i] - \ip{a}{\psi_i} = - \int_{\theta_i} (\bu|_{s^h} \cdot \bn_{s^h}- a)\, \psi_i  \label{eq:rhozero}
\end{equation}
and
\begin{align}
\rho_i'(s^h; 0) &= (F^h)'(s^h)[\psi_i,\psi_i]  \label{eq:rholocalderiv} \\
  &= - \int_{\theta_i} \lim_{\eps\to 0^+} \frac{(\bu|_{s^h+\eps\psi_i} - \bu|_{s^h}) \cdot \bn_{s^h+\eps\psi_i}}{\eps} \psi_i - \int_{{\hat\theta}_i} \lim_{\eps\to 0^+} \frac{\bu|_{s^h+\eps\psi_i} \cdot \bn_{s^h+\eps\psi_i}}{\eps} \psi_i.  \notag
\end{align}
